### 2018-08-08 paper/finitecomplete-failures.tex
Theorem on trace refinement testing added
Fixes after suggestions by Wen-ling

If $P/s$ is the STOP-process, this is revealed by its initials being empty. In this case, the test may terminate 
successfully (second operand of the external choice in $UF(p,s)$). Note that at the same time, any (now illegal) event of the alphabet is also accepted by the test.
Therefore,
if the SUT would still accept an event in a state where $P/s$ is supposed to have stopped, there exists a test execution which terminates with FAIL by choosing the first
operand of the external choice.

If the length of $s$ is still less than $p$, the test accepts any event from the initials
$[P/s]^0$ and continues recursively as sub-process $U(p,s.e)$ (third operand of the external choice). A test of this type is called \emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts its consecutive behaviour to the event selected by the SUT.

@@

\begin{theorem}
Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ 
whose normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal D}$ as
the set of all divergence-free CSP processes over alphabet $\Sigma$, whose transition graph
has at most $q$ states with $q \ge p$. 
Then the test suite 
\[
\TS_F = \{ U_F(k)~|~0 \le k < pq  \}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
\begin{proof}
To prove soundness of $\TS_F$, suppose that $Q\in{\cal D} \wedge P\lessdet_F Q$. Then
$\trc(Q)\subseteq \trc(P)$, and Lemma~\ref{lemma:hseta} implies that 
for all traces $s$ of $Q$, every $H\in\text{minHit}(P/s)$ is a 
    hitting set for $\minaccs(Q/s)$. As a consequence, when running in parallel
with $Q$, any adaptive test $U_F(p)$ will always enter the branches 
(\ref{eq:ufb}), (\ref{eq:ufc}),  or (\ref{eq:ufd}) of the external choice construction
for $U_F(p,s)$. Branch (\ref{eq:ufb}) always leads to a PASS verdict, branch (\ref{eq:ufc})
always to test continuation without a verdict. For the last branch, we note that 
any selected minimal hitting set $H\in\text{minHit}(P/s)$ has a non-empty intersection with
each of the minimal acceptances of $Q/s$. As a consequence, $Q/s$ will never block when offered events from $H$, and the test terminates with PASS event $\epass$. Note that this argument requires that $Q$ is free of livelocks, because otherwise the PASS-events might not become visible, due to unbounded sequences of hidden events performed by $Q$. Note further, that this proof did not refer in any way to the size of $Q$'s normalised transition graph, so the test suite is sound for {\it all} non-divergent CSP process refining $P$.

To prove exhaustiveness, consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. This non-conformance can be caused in two possible ways.
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that 
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation} 
\end{description}
It has to be shown for each of the two possibilities that at least one test execution of some $(Q\parallel[\Sigma] U(k))$ with $k < pq$ ends with the FAIL event $\efail$.

For the first case, consider a  trace $s.e\in\trc(Q)$ such that $s\in\trc(P)$, but 
$s.e\not\in\trc(P)$. Then $s$ is also a trace of the product graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}. From the construction of $G$ described there, we know that $G$ has at most $pq$ reachable states, because $G(P)$ has $p$ states, and $G(Q)$ has at most $q$ states. Suppose that $G/s = (n_P,n_Q)$. Applying Lemma~\ref{lemma:extendV} implies that this state can be reached by a trace $u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition function of $G$ implies that $u$ is also a trace of $P$ and $Q$. Since test $U_F(pq-1)$ accepts all traces of $P$ up to length $pq$, $u$ is also a trace of this test, and, by construction, $U_F(pq-1)/u = U_F(pq-1,u)$. Since $s.e\not\in\trc(P)$,
$e$ is an element of $\Sigma-[P/u]^0$. Therefore, in at least one execution, $U_F(pq-1,u)$ executes its first branch (\ref{eq:ufa}) with this event $e$, so that the test fails with event $\efail$. Again, the assumption of non-divergence of Q is needed for this conclusion.

>>>>

If $P/s$ is the STOP-process, this is revealed by its initials being empty. In this case, the test may terminate 
successfully (second operand of the external choice in $U_F(p,s)$). Note that at the same time, any (now illegal) event of the alphabet is also accepted by the test.
Therefore,
if the SUT would still accept an event in a state where $P/s$ is supposed to have stopped, there exists a test execution which terminates with FAIL by choosing the first
operand of the external choice.

If the length of $s$ is still less than $p$, the test accepts any event from the initials
$[P/s]^0$ and continues recursively as sub-process $U_F(p,s.e)$ (third operand of the external choice). A test of this type is called \emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts its consecutive behaviour to the event selected by the SUT.

@@

\begin{theorem}\label{th:failurestest}
Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ 
whose normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal D}$ as
the set of all divergence-free CSP processes over alphabet $\Sigma$, whose transition graph
has at most $q$ states with $q \ge p$. 
Then the test suite 
\[
\TS_F = \{ U_F(k)~|~0 \le k < pq  \}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
\begin{proof}
To prove soundness of $\TS_F$, suppose that $Q\in{\cal D} \wedge P\lessdet_F Q$. Then
$\trc(Q)\subseteq \trc(P)$, and Lemma~\ref{lemma:hseta} implies that 
for all traces $s$ of $Q$, every $H\in\text{minHit}(P/s)$ is a 
    hitting set for $\minaccs(Q/s)$. As a consequence, when running in parallel
with $Q$, any adaptive test $U_F(p)$ will always enter the branches 
(\ref{eq:ufb}), (\ref{eq:ufc}),  or (\ref{eq:ufd}) of the external choice construction
for $U_F(p,s)$. Branch (\ref{eq:ufb}) always leads to a PASS verdict, branch (\ref{eq:ufc})
always to test continuation without a verdict. For the last branch, we note that 
any selected minimal hitting set $H\in\text{minHit}(P/s)$ has a non-empty intersection with
each of the minimal acceptances of $Q/s$. As a consequence, $Q/s$ will never block when offered events from $H$, and the test terminates with PASS event $\epass$. Note that this argument requires that $Q$ is free of livelocks, because otherwise the PASS-events might not become visible, due to unbounded sequences of hidden events performed by $Q$. Note further, that this proof did not refer in any way to the size of $Q$'s normalised transition graph, so the test suite is sound for {\it all} non-divergent CSP process refining $P$.

To prove exhaustiveness, consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. This non-conformance can be caused in two possible ways.
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that 
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation} 
\end{description}
It has to be shown for each of the two possibilities that at least one test execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the FAIL event $\efail$ or without giving any verdict. The latter case is also interpreted   as FAIL, since then  
$(\epass\then\Stop)$ is no longer refined by the test execution in the failures model.

For the first case, consider a  trace $s.e\in\trc(Q)$ such that $s\in\trc(P)$, but 
$s.e\not\in\trc(P)$. Then $s$ is also a trace of the product graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}. From the construction of $G$ described there, we know that $G$ has at most $pq$ reachable states, because $G(P)$ has $p$ states, and $G(Q)$ has at most $q$ states. Suppose that $G/s = (n_P,n_Q)$. Applying Lemma~\ref{lemma:extendV} implies that this state can be reached by a trace $u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition function of $G$ implies that $u$ is also a trace of $P$ and $Q$. Since test $U_F(pq-1)$ accepts all traces of $P$ up to length $pq-1$, $u$ is also a trace of this test, and, by construction, $U_F(pq-1)/u = U_F(pq-1,u)$. Since $s.e\not\in\trc(P)$,
$e$ is an element of $\Sigma-[P/u]^0$. Therefore, in at least one execution, $U_F(pq-1,u)$ executes its first branch (\ref{eq:ufa}) with this event $e$, so that the test fails with event $\efail$. Again, the assumption of non-divergence of Q is needed for this conclusion.
